---
title: "Preparing Data for Constructing Diversification Measures (eg, MS = 1 if a multiple-segment firm)"
output:
  html_notebook: default
  # html_document: default
  # word_document: default
---


# Load raw data and required library
```{r}

# library(magrittr)
library(tidyverse)
library(lubridate)
library(stargazer)


name_df = "segmentHist"

mainDir = '' #Sys.getenv("Data") #'D:/github/Data/'
subDir = ''  #'/Segment/'
#dir.create(file.path(mainDir, subDir), showWarnings = FALSE, recursive = TRUE)


###################################################################################
#---time_stamp = "2019-10-31_23.55.19"
# To review the data structure of the raw data:
rawData <- read.csv(paste0(mainDir, subDir, 'compsegd_1990_1991.csv')) %>% 
    mutate(
      datadate_int = datadate, # keep a copy of the original integer format
      srcdate_int = srcdate,
      datadate = ymd(datadate), # convert to Date format
      srcdate = ymd(srcdate),
      )
# change all column variable names to lowercase
names(rawData) <- tolower(names(rawData))
#---rawData <- readRDS(paste0(mainDir, subDir, name_df, "_raw", time_stamp, '.rds')) 
###################################################################################


earliest = "1990-01-01" #"1998-01-01" 
lastdate = "1999-12-31"    #"1999-12-31"
rawData %>% 
  filter(datadate >= earliest & datadate <= lastdate) %>% 
  saveRDS(paste0(mainDir, subDir, name_df, "_raw", 'Example.rds'))


#=====================================================
# segment Variable list for this dataset
segmvars_list = "datadate, srcdate, srcs, upds, gvkey, sics1, sics2, sics3, 
stype, sid, snms, soptp1, soptp2, geotp, naicsh, naicss1, naicss2, naicss3,

emps, emps_dc, emps_fn, obs, obs_dc, 
intseg, intseg_dc, oelim, oelim_dc, 
esubs, esubs_dc, esubs_fn, ivaeqs, ivaeqs_dc, 
salexg, salexg_dc, spis, spis_dc, xidos, xidos_dc, 

atlls, atlls_dc, capxs, capxs_dc, capxs_fn, caxts, caxts_dc, 
cogss, cogss_dc, dps, dps_dc, gdwls, gdwls_dc, 

ias, ias_dc, ibs, ibs_dc, iints, iints_dc, 
nis, nis_dc, nopxs, nopxs_dc, nxints, nxints_dc, ocaxs, ocaxs_dc, 
oiadps, oiadps_dc, oibdps, oibdps_dc, ops, ops_dc, ops_fn, 
ppents, ppents_dc, ptis, ptis_dc, rds, rds_dc, rds_fn, 

revts, revts_dc, sales, sales_dc, sales_fn,  
txts, txts_dc, txws, txws_dc, 
xints, xints_dc, xsgas, xsgas_dc, 

curcds, 
isosrc 
"  

#=====================================================
# firm Variable list for this dataset
firmvars_list = "datadate gvkey 
fic adrr
fyear sich at sale oiadp ceq prcc_f csho capx  gdwl" %>% 
  str_split("[ |\n]+", simplify = T) %>% c()  # "[ |\n]+" matches any number of spaces or line breaks (\n)

```


# Data integrity checks: Remove problematic observations (eg, gvkey-datadate obs. with missing-value sics1, etc)
```{r}

earliest = "1990-01-01" #"1998-01-01" 
lastdate = "1999-12-31"    #"1999-12-31"

data_cleaned <- rawData %>% 
  filter(srcdate >= datadate) %>% # remove at least 6 data error cases
################################################################## <-----------------------------!!!  
# What if Custodio forgot to impose this ???  
  filter(srcdate == datadate) %>% # confine to non-restated data
################################################################## 
  # extract year, month, day info from datadate and srcdate to facilitate data integrity checks to be done later
  mutate(dyear = year(datadate), dmonth = month(datadate), dday = day(datadate),
         syear = year(srcdate), smonth = month(srcdate), sday = day(srcdate),
         ) %>% 
  # define for segment data the equivalent of Compustat's fyear for later merging with Fundamental annual data
  mutate(
    fyears = case_when(
      dmonth <= 5 ~ year(datadate) - 1,
      TRUE ~ year(datadate),
      ),
    ) %>% 
  #=========================================================
  # upds = 2 means newswire, incomplete, trading suspended, etc; = 1 means only market data, csho 
  filter(upds == 3 & !is.na(upds)) %>% # = 3 means final
  filter(datadate >= earliest & datadate <= lastdate) %>% 
############################################################### <----------------------------------------- !!!  
  filter(stype == "BUSSEG") %>% # keep only BUSSEG
  mutate(rm_flag = FALSE) %>% # set initial value of rm_flag (ie, by default, do not remove)
################################################################################################
  mutate(snms = tolower(snms)) %>% # standardize segment name to lowercase
  mutate_at(vars(starts_with("sics")), ~ as.integer(.)) %>%  # first, convert sics's to integer type to be consistent with sich
  # Create sics_my to allow for the possibility of later classifying undesirable cases into
  #   negative-value categories (eg, -6666 for "undesirable" financial segment)
  mutate(sics_my = sics1) %>% 
################################################################################################
                  # select(datadate, srcdate, srcs, upds, fyears, gvkey, 
                  #        rm_flag, sics_my, sics1, sics2, sics3, snms, soptp1, soptp2, geotp, 
                  #        everything()) %>% 
  select(everything())  


############################################################
#!!! Note: types of segments (soptp1, soptp2, geotp) are important
# 
# soptp1 = "PD_SRVC" "GEO"  "MARKET" "DIV" "OPER" NA
# soptp1 = "PD_SRVC" "GEO"  "MARKET" "DIV" NA
# geotp = "3" "2" "1" NA
############################################################

```


# Data integrity checks: Prelim look of categorical data and summary of continuous variables
```{r}

facVars <- "dmonth, smonth, dday, sday, srcs, upds, stype, sid, soptp1, soptp2, geotp, curcds, isosrc" %>%  
  str_split(", ", simplify = TRUE) %>% 
  c()

# Prelim look of various categorical variables
data_cleaned %>% 
  # turn the following into factors only for the purpose of examining the categorical and continuous variables
  mutate_at(vars(facVars), ~ as.factor(.)) %>% 
  select(datadate, srcdate, fyears, dyear, syear, dmonth, smonth, dday, sday, 
         srcs, upds, stype, sid, 
         rm_flag, soptp1, soptp2, geotp, curcds, isosrc
         ) %>% 
  summary()


cat("\n\nsummary of continuous variables:\n\n")
data_cleaned %>%
  select(-facVars) %>% 
  select_if(is.numeric) %>%
  summary() %>% 
  print()

```


# Data integrity checks: Confirm no duplicate sid for each gvkey-datadate (or gvkey-dyear-dmonth)
```{r}
#rm(rawData0)

# no duplicate for each (gvkey, datadate, sid)
data_cleaned %>% 
  group_by(gvkey, datadate, sid) %>% 
  filter(n() > 1) %>% 
  nrow() 
# Expecting 0 found 


# no duplicate for each (gvkey, dyear, dmonth, sid)
data_cleaned %>% 
  group_by(gvkey, dyear, dmonth, sid) %>% 
  filter(n() > 1) %>% 
  nrow() 
# Expecting 0 found 


#==========================================================
# duplicates for each (gvkey, [dyear], sid) due to change of FYE or segment name/structure
data_cleaned %>% 
  group_by(
    stype,  
    gvkey, dyear, sid) %>% 
  filter(n() > 2) %>%
  nrow()
# Expecting 0 found for n() > 2 (assuming n() = 2 caused by change of FYE only)
        # arrange(gvkey, stype, sid, datadate) %>% 
        # View()

```


# Save cleaned data for further processing and regression analysis
```{r}

saveRDS(data_cleaned, file = paste0(mainDir, subDir, name_df, "_clean", 'Example.rds'))

```



